---
title: "Untitled"
author: "Jiangran Liu"
date: "3/2/2020"
output: md_document
---

```{r}
## loading packages
library(lubridate)
library(dplyr)
library(tidyr)
library(nycflights13)
library(ggplot2)
```

## Make a plot with three facets, one for each airport in the weather data. The x-axis should be the day of the year (1:365) and the y-axis should be the mean temperature recorded on that day, at that airport.
```{r}
weather %>% mutate(day = yday(time_hour)) %>% 
  group_by(origin, day) %>% 
  summarize(temp = mean(temp, na.rm=T)) %>% 
  ggplot(aes(x = day, y = temp)) +
  geom_line()+facet_wrap(~origin)
```

## Make a non-tidy matrix of that data where each row is an airport and each column is a day of the year.
```{r}
weather %>% mutate(day = yday(time_hour)) %>% 
  group_by(origin, day) %>% 
  summarize(temp = mean(temp, na.rm=T)) %>% 
  pivot_wider(names_from=day, values_from = temp)
```

## For each (airport, day) contruct a tidy data set of the airport’s “performance” as the proportion of flights that departed less than an hour late.
```{R}
perf = flights %>% mutate(day = yday(time_hour)) %>% 
  group_by(origin, day) %>% 
  summarize(performance = mean(dep_delay<60, na.rm = T))
```

## Construct a tidy data set to that give weather summaries for each (airport, day). Use the total precipitation, minimum visibility, maximum wind_gust, and average wind_speed.
```{R}
weat = weather %>% mutate(day = yday(time_hour)) %>% 
  group_by(origin, day) %>% 
  summarize(precipitation = sum(precip, na.rm=T), 
            visibility = min(visib, na.rm=T), 
            windMean = mean(wind_speed, na.rm=T)) 
```

## Construct a linear model to predict the performance of each (airport,day) using the weather summaries and a “fixed effect” for each airport. Display the summaries.
```{R}
data = left_join(perf, weat)
lm = lm(performance ~ origin+precipitation+visibility+windMean, data = data)
summary(lm)
```

## Repeat the above, but only for EWR. Obviously, exclude the fixed effect for each airport.
```{R}
EWR_data = filter(data, origin == "EWR")
EWR_lm = lm(performance ~ precipitation+visibility+windMean, data = EWR_data)
summary(EWR_lm)
```